Evolutionary market ecologies and market malfunction\\ Draft

### 2021-06-29 draft.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\code}{\texttt}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{bibliography.bib}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms 
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx 
\newcommand{\var}[1]{{\ttfamily#1}}% variable

\title{Evolutionary market ecologies \\ Model \& Numerical experiments}
\author{Aymeric Vi\'{e}}
\date{\today}

@@

\section{Trend following Model description}

\subsection{Initialisation}

@@

Where $g$ is the dividend growth rate, $\sigma$ its variance, $\omega$ is the autocorrelation parameter of the process, and $Z$ is a standard Wiener process. Based on estimates from Lebaron \cite{lebaron2001empirical} and Scholl et al. \cite{scholl2020market}, we take $g=2\%$ annually and $\sigma = 6\%$ annually. \par
The market is initialised with an initial price of $100$ that matches the fundamental value of the asset, and initial dividends of $100$.

\subsection{Strategy and market behavior}
At each time period, trend followers look at past prices to decide the content of their next portfolio, choosing between cash and an asset. The share of their wealth that will be devoted to asset shares is proportional to the agent trading signal $\phi(t)$. Here, the agent compares the logarithm of the prices at last period, and at another period $\theta$. This $\theta$ is the time horizon of this agent's trend following strategy, and is subject to evolution. We consider $\theta \in (1, H]$ with $H$ the maximum allowed time horizon. In \cite{scholl2020market}, trend followers all had $\theta = 2$; this generalisation allows us to look at a larger space of strategies.
% here we write what the agents do, trading signal (write an example with a time series and a few different agents? that would be very nice. Inspire from David's figure 1?), excess demand, market clearing
% mention the next strategy extensions
% mention hypermutation & replacement

@@

@@

Agent $i$'s wealth $W_i(t)$ evaluated at time $t$ is computed with:
\begin{equation}
    W_i(t) = C_i(t) + S_i(t)p(t) - L_i(t)
\end{equation}

\subsection{Market clearing}

@@

@@

The non-lerning case (replacement dynamics only) can be illustrated with:
\begin{itemize}
    \item No selection (selection rate equal to 0)
    \item No crossover (crossover rate equal to 0)
    \item No mutation (mutation rate equal to 0)
\end{itemize}
The objective is to assess the behavior of the model in this no-learning setup, establish a benchmark.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\newcommand{\code}{\texttt}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{bibliography.bib}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{algorithm}
\usepackage{algpseudocode}
\newcommand{\var}[1]{{\ttfamily#1}}
\newcommand\boldred[1]{\textcolor{red}{\textbf{#1}}}

\title{Evolutionary market ecologies and market malfunction\\
Draft}
\author{Aymeric Vi\'{e}}
\date{\today}

@@

\begin{abstract}
    \boldred{Abstract TBA}
\end{abstract}

\section{Introduction}
\boldred{Introduction TBA}
\boldred{Everything related to GP and multi asset: in footnotes}
% Maybe paragraphs could be
% market ecology
% evolutionary computation
% evolution in financial markets

\section{Model}

\subsection{Initialisation}
\boldred{To update with new structure}

@@

Where $g$ is the dividend growth rate, $\sigma$ its variance, $\omega$ is the autocorrelation parameter of the process, and $Z$ is a standard Wiener process. Based on estimates from Lebaron \cite{lebaron2001empirical} and Scholl et al. \cite{scholl2020market}, we take $g=2\%$ annually and $\sigma = 6\%$ annually. \par
The market is initialised with an initial price of $100$ that matches the fundamental value of the asset, and initial dividends of $100$. 
\boldred{Add a note for the multi asset thing. We actually need to think about multiple assets now that we are doing these big changes. Notably, the ED and TS matrix would change to be $n\timesK$ matrices, where n pop size and K number of assets. The agent strategies change as well to be a K-vector of strategies. It would be better to have a very general case, a very general structure, and so that we have the right structure and only need particular settings or specific small fixes to make it work and extend it.}

\subsection{Strategy and market behavior}
\boldred{Update for the other strategies as well}
At each time period, trend followers look at past prices to decide the content of their next portfolio, choosing between cash and an asset. The share of their wealth that will be devoted to asset shares is proportional to the agent trading signal $\phi(t)$. Here, the agent compares the logarithm of the prices at last period, and at another period $\theta$. This $\theta$ is the time horizon of this agent's trend following strategy, and is subject to evolution. We consider $\theta \in (1, H]$ with $H$ the maximum allowed time horizon. In \cite{scholl2020market}, trend followers all had $\theta = 2$; this generalisation allows us to look at a larger space of strategies.
% here we write what the agents do, trading signal (write an example with a time series and a few different agents? that would be very nice. Inspire from David's figure 1?), excess demand, market clearing
% mention the next strategy extensions
% mention hypermutation & replacement

@@

\boldred{With multiple assets, how do we formulate the excess demand? Replace E and TS and p and S by the matrix form (generalised ED function)?}

@@

Agent $i$'s wealth $W_i(t)$ evaluated at time $t$ is computed with:
\begin{equation}
    W_i(t) = C_i(t) + S_i(t)p(t) - L_i(t)
\end{equation}
\boldred{Extends to a vector S p with multiple assets}

\subsection{Market clearing}
\boldred{Add something to imagine clearing with a multi asset environment, just consists in adding all the assets with each its own price}

@@

We impose a non-negativity constraint on the cash. 
% In other words, the agents cannot enter into cash debt to buy more assets. 
This must work with leverage.

@@

The non-learning case (replacement dynamics only) can be illustrated with:
\begin{itemize}
    \item No selection (selection rate equal to 0)
    \item No crossover (crossover rate equal to 0)
    \item No mutation (mutation rate equal to 0)
\end{itemize}
The objective is to assess the behavior of the model in this no-learning setup, establish a benchmark.
